\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2013-09-04 paper.tex
Rephrasing, fixing fixmes...

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{spark-oro}
    \caption{Functional overview of knowledge base ({\sc Oro} server, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
        \label{fig|spark-oro}
\end{figure}

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic interpretation. We rely on a dedicated geometric and temporal
reasoning module called SPARK (\emph{SPAtial Reasoning \&
Knowledge}~\cite{Sisbot2011}). It is a situation assessment reasoner that
generates symbolic knowledge from the geometry of the environment with respect
to relations between objects, robots and humans (Fig.~\ref{fig|spark-oro}),
also taking into account the different perspective that each agent has on the
environment.

SPARK is an \emph{amodal}~\cite{Mavridis2006} geometric model of the
environment that serves both as basis for the fusion of the perception
modalities and as bridge with the symbolic layer. This geometric model is
continuously updated at run-time by the robot based on its sensors (in our
experiments, objects are identified and localised through 2D barcodes, while
humans are tracked with Kinect-like devices, optionally assisted by motion
capture to accurately track the head motion, which is required to compute what
the human is looking at).

\fxnote{Add 2-3 lines to summarize the AI highlights of SPARK}

@@

SPARK currently provides a simple implementation of such a functionality. The
only managed hypotheses are \emph{in container} and \emph{in agent hand}. We
can have only one hypothesis at the same time. Hypothesis validity is checked
geometrically in case of incoming perception values.

\paragraph{Primitive action recognition} Monitoring human activity is crucial
to maintain a coherent state of the world. Full human action and activity
monitoring is a task that requires knowledge and reasoning both on high level
facts like goals, intentions and plans, as well as bottom-up data from agent
and object motions. Simple temporal and geometric reasoning on human hand
trajectories and potential objects placements can provide some useful clues for
high level human monitoring processes. We call this temporal and geometric
reasoning \emph{primitive action recognition}.

For example, a \emph{pick}, a \emph{throw} or a \emph{place} action
can be recognized by observing that an object on table and an empty
human hand are close to each other, or that the human hand holding an
object is close to a container, etc. Human hand position is either
directly perceived or inferred from its initial perceived trajectory\fxwarning{we really do that?}.
We have a simple implementation of such a primitive action recognition
in SPARK that relies on monitoring human hand and its motion near
objects or above containers.

@@

SPARK enables \emph{perspective-taking}: spatial relations between entities can
be computed from different viewpoints, which let the robot build for each
agents it interacts with a different, \emph{perspective-aware} symbolic model
of the environment. These models are separately stored in the knowledge base.
\fxwarning{Rephrase: already mentioned before}

@@

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. We use in our architecture the
HATP planner (for Human Aware Task Planner~\cite{Alili2008, Alili2009}).  HATP
is based on a Hierarchical Task Network (HTN) refinement, which performs an
iterative task decomposition into sub-tasks until reaching atomic
actions\fxwarning{Rephrase: the definition is actually generic to HTN, not HATP
specfic}. The \emph{planning domain} defines the set of methods describing how
to decompose a task and represents the procedural knowledge of the robot. It is
stored outside of the central knowledge base.

HATP is able to produce plans for the robot's actions as well as for the other
participants (humans or robots). It can be tuned by setting up different costs
depending on the actions to apply and by taking into account a set of
constraints called social rules. This tuning aims at adapting the robot's
behavior according to the desired level of cooperation of the robot.

@@

\paragraph{Context Management Policies} \fxfatal{Rewrite this section to level up AI aspects (Mathieu)}
Building, updating and maintaining a correct state of the world at
geometric and symbolic level is crucial to the capacity of the robot
to carry on successfully a multi-step interaction with a human. Tight
integration between the robot controller and the geometric and
temporal reasoning functions in SPARK and symbolic facts and beliefs
management in ORO is central.

@@

Robot must reacts to change in the world not linked to robot action to drive
world update. Robot controller uses SPARK to monitor human hand motion and
primitive action recognition for \emph{pick} and \emph{throw}. As mentioned
above these primitive action recognition should be used with higher level
information on goals, intentions and plans and some exploration to achieve
complex human action and activity monitoring.  In the current implementation,
these primitive actions are over-optimistically interpreted as the
corresponding actions \emph{pick} object and \emph{throw} object.

@@

    \caption{Snapshot of the film of the ``Clean the Table'' scenario. The
    physical situation, the SPARK model, and the current step of the plan can
    be seen.}

@@

This last section goes over again the main contributions we have presented
so far: why are they relevant to robotics and artificial intelligence? How do
they compare to similar approaches? How do we justify our technical choices and
what remain to be developed/explored in the future?

@@

One of the main limit of our current approach is the way semantics are
conveyed: because we use 2D markers for artifacts, we skip most of the object
extraction and recognition challenges. Each object has a unique identifier,
which enable us to load a proper CAD model suitable for 3D spatial reasoning
and prevent recognition mistakes in the knowledge base. While {\sc Spark}
algorithms themselves are ignorant of the input sources, and would work equally
well with a full object recognition stack, it is not currently supported.

\fxwarning{spatial reasoning -> mention we do not support extensive temporal reasoning}

@@

Altogether, the components we have presented compose an architecture that we
call \emph{knowledge-oriented}:

@@

Those items are not new \emph{per se}. It seems however interesting to
underline the shift of focus this implies during the design and integration
phases: components of our deliberative layer are defined and glued together by
the knowledge they produce and consume.

@@

Robotics is traditionally regarded as the prototypical testbed for \emph{embodied}
artificial intelligence, and this dimension is especially prevalent in
human-robot interaction, where agents have to share a joint physical
environment.

>>>>

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{spark-oro}
    \caption{Functional overview of knowledge base ({\sc Oro} server, top part) and the geometric situation assessment module (\emph{{\sc Spark}}, bottom part)}
        \label{fig|spark-oro}
\end{figure}

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic interpretation. We rely on a dedicated geometric and temporal
reasoning module called {\sc Spark} (\emph{SPAtial Reasoning \&
Knowledge}~\cite{Sisbot2011}). It is a situation assessment reasoner that
generates symbolic knowledge from the geometry of the environment with respect
to relations between objects, robots and humans (Fig.~\ref{fig|spark-oro}),
also taking into account the different perspective that each agent has on the
environment.

{\sc Spark} is an \emph{amodal}~\cite{Mavridis2006} geometric model of the
environment that serves both as basis for the fusion of the perception
modalities and as bridge with the symbolic layer. This geometric model is
continuously updated at run-time by the robot based on its sensors (in our
experiments, objects are identified and localised through 2D barcodes, while
humans are tracked with Kinect-like devices, optionally assisted by motion
capture to accurately track the head motion, which is required to compute what
the human is looking at).

\fxnote{Add 2-3 lines to summarize the AI highlights of {\sc Spark}}

@@

{\sc Spark} currently provides a simple implementation of such a functionality. The
only managed hypotheses are \emph{in container} and \emph{in agent hand}. We
can have only one hypothesis at the same time. Hypothesis validity is checked
geometrically in case of incoming perception values.

\paragraph{Primitive action recognition} Monitoring human activity is crucial
to maintain a coherent state of the world. Full human action and activity
monitoring is a task that requires knowledge and reasoning both on high level
facts like goals, intentions and plans, as well as bottom-up data from agent
and object motions. Simple temporal and geometric reasoning on human hand
trajectories and potential objects placements can provide some useful clues for
high level human monitoring processes. We call this temporal and geometric
reasoning \emph{primitive action recognition}. Those primitive actions are
assessed through monitoring situations like ``an empty hand is close to an
object on a table'' (precursor for a \emph{pick}), or ``a hand holding an
object is over a container'' (precursor for a \emph{throw}).

@@

{\sc Spark} enables \emph{perspective-taking}: spatial relations between entities can
be computed from different viewpoints, which let the robot build for each
agents it interacts with a different, \emph{perspective-aware} symbolic model
of the environment. These models are separately stored in the knowledge base.
\fxwarning{Rephrase: already mentioned before}

@@

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. We use in our architecture the
HATP planner (for Human Aware Task Planner~\cite{Alili2008, Alili2009}).  HATP
is a refinement of a Hierarchical Task Network (HTN). As an HTN, it performs an
iterative task decomposition into sub-tasks until reaching atomic actions. The
\emph{planning domain} defines the set of methods describing how to decompose a
task and represents the procedural knowledge of the robot. It is stored outside
of the central knowledge base, using a specific formalism.

HATP originality resides in its ability to produce plans for the robot's
actions \emph{as well as for the other participants} (humans or robots): the
robot plans for itself, but also for its interactors, which effectively endows the robot
with new interaction capabilities: It can partially predict the behaviour, and
thus monitor the engagement, of its human partner; the planner generates
synchronisation points (arrow going from the robot stram to the human stream on
Figure~\ref{plan_hatp1}) between the agents; the controller can also verbalize
the plans to explain its partner how a task may be shared~\cite{warnier2012when}.

HATP can be tuned by setting up different costs depending on the actions to
apply and by taking into account a set of constraints called \emph{social
rules}. This tuning aims at adapting the robot's behavior according to the
desired level of cooperation of the robot.

@@

\paragraph{Context Management Policies} \fxfatal{Rewrite this section to level up AI aspects (Mathieu)}
Building, updating and maintaining a correct state of the world at
geometric and symbolic level is crucial to the capacity of the robot
to carry on successfully a multi-step interaction with a human. Tight
integration between the robot controller and the geometric and
temporal reasoning functions in {\sc Spark} and symbolic facts and beliefs
management in ORO is central.

@@

Robot must reacts to change in the world not linked to robot action to drive
world update. Robot controller uses {\sc Spark} to monitor human hand motion and
primitive action recognition for \emph{pick} and \emph{throw}. As mentioned
above these primitive action recognition should be used with higher level
information on goals, intentions and plans and some exploration to achieve
complex human action and activity monitoring.  In the current implementation,
these primitive actions are over-optimistically interpreted as the
corresponding actions \emph{pick} object and \emph{throw} object.

@@

    \caption{Snapshot of the film of the ``Clean the Table'' scenario. The
    physical situation, the {\sc Spark} model, and the current step of the plan can
    be seen.}

@@

This last section goes over again the main contributions we have presented
so far: why they are relevant to robotics and artificial intelligence, how do
they compare to similar approaches, how do we support our technical choices and
what remains to be developed/explored in the future.

@@

One of the main limit of our current approach is the way semantics are
conveyed: because we use 2D markers for artifacts, we skip most of the object
extraction and recognition challenges. Each object has a unique identifier,
which enable us to load a proper CAD model suitable for 3D spatial reasoning
and prevent recognition mistakes in the knowledge base. While {\sc Spark}
algorithms themselves are ignorant of the input sources, and would work equally
well with a full object recognition stack, we have not tackled it yet.

Besides, temporal reasoning (essential for accurate action recognition for
instance) is not properly addressed in our stack. Temporal reasoning is used
only locally, and does not allow for tracking of long or global events.

@@

Altogether, the components we have presented compose a
\emph{knowledge-oriented} architecture:

@@

Those items are not new \emph{per se}. It is however interesting to
underline the shift of focus this implies during the design and integration
phases of robots: components of our deliberative layer are defined and glued together by
the knowledge they produce and consume.

It reveals how the robotic community is ``catching up'' with semantic
technologies. Human-robot interaction, because it supposes operations in
environments with complex semantics, drives this evolution.

@@

Robotics is traditionally regarded as the prototypical instance of \emph{embodied}
artificial intelligence, and this dimension is especially prevalent in
human-robot interaction, where agents have to share a joint physical
environment.
